---
title: "probabilities"
author: "John Morse"
date: "3/24/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Probabilities

_NOTE: All Page references are to the R Cookbook_

Suppose proportion of people with red hair is 23%.
If we select a sample of 38 people:
- What is probability of exactly 4 people with red hair?
- on the TI-84: binompdf(38, .23, 4) = 0.0286
Page 223, density function with binomial distribution

```{r}
dbinom(4, size = 38, prob = 0.23)
```

- What is the probability at least 4 people have red hair?
- on the TI-84: 1 - binomcdf(38, .23, 3) = 0.9854
Page 224, distribution function with binomial distribution

```{r}
1 - pbinom(3, size = 38, prob = 0.23)
```

Alternatively:
```{r}
pbinom(3, size = 38, prob = 0.23, lower.tail = FALSE)
```

Note the definition of lower.tail:
- if TRUE (default), probabilities are P[X ≤ x], otherwise, P[X > x]
- setting to FALSE, it is getting GREATER THAN x, NOT greater than or equal

#Problem 2
Average score in class is 67.4, standard deviation is 11.3. Exam scores are normally distributed.
	- If one student is selected at random, what is the probability the student will score less than 60.5?
		○ In this case, we are selecting an individual from the population, so we use a simple z score:
		○ z=(x−μ)/σ
		○ x = 60.5, mean = 67.4, std dev = 11.3
		○ z = -0.6106
		○ normalcdf(-1E99, -0.6106, 0, 1) = 0.2707
Page 243 and page 225

```{r}
# Calculating a z score
z <- (60.5 - 67.4) / 11.3
print(z)

#Using the z-score to get the probability
pnorm(z, mean = 0, sd = 1, lower.tail = TRUE)

```

		
		
		
	- If a group of 10 students is selected, find probability that their mean score would be greater than 60.5?
		○ In this case, we are selecting a sample mean
		○ z=(x ̅−μ)/(σ/√n)
		○ n = 10
		○ z = -1.9309
		○ 1 - normalcdf(-1E99, -1.9309, 0, 1) = 0.9733
	- The Top 13% of all students score higher than ???
		○ Use page 246
		○ Since this is the Top 13%, 1 - 0.13 = 0.87
		○ Find the z-score corresponding to the area 0.87
		○ invNorm(0.87, 0, 1, LEFT) = 1.1264
		○ Use the z-score and the formula:
			○ x=μ+(z ∗ σ)
		○ X = 67.4 + (1.1264 * 11.3) = 80.1283 






