---
title: "probabilities"
author: "John Morse"
date: "3/24/2020"
output: 
  
  powerpoint_presentation: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Probabilities

**NOTE: All Page references are to the R Cookbook**

## Problem 1

Suppose proportion of people with red hair is 23%.
If we select a sample of 38 people:

* What is probability of exactly 4 people with red hair?
* on the TI-84: binompdf(38, .23, 4) = 0.0286

*Refer to: Page 223, density function with binomial distribution*

```{r}
dbinom(4, size = 38, prob = 0.23)
```

* What is the probability at least 4 people have red hair?
* On the TI-84: 1 - binomcdf(38, .23, 3) = 0.9854

*Refer to: Page 224, distribution function with binomial distribution*

```{r}
1 - pbinom(3, size = 38, prob = 0.23)
```

Alternatively:
```{r}
pbinom(3, size = 38, prob = 0.23, lower.tail = FALSE)
```

Note the definition of lower.tail:

* if TRUE (default), probabilities are P[X ≤ x], otherwise, P[X > x]
* setting to FALSE, it is getting GREATER THAN x, NOT greater than or equal

## Problem 2

Average score in class is 67.4, standard deviation is 11.3. Exam scores are normally distributed.

```{r}
# Set up primary values
mu <- 67.4
sdev <- 11.3
```

* If one student is selected at random, what is the probability the student will score less than 60.5?
    * In this case, we are selecting an individual from the population, so we use a simple z score:
    * z=(x−μ)/σ
    * x = 60.5, mean = 67.4, std dev = 11.3
    * z = -0.6106
    * On the TI-84: normalcdf(-1E99, -0.6106, 0, 1) = 0.2707

*Refer to: Page 243 and page 225*

```{r}
# Calculating a z score
z <- (60.5 - mu) / sdev
print(z)

#Using the z-score to get the probability
pnorm(z, mean = 0, sd = 1, lower.tail = TRUE)

```

* If a group of 10 students is selected, find probability that their mean score would be greater than 60.5?
    * In this case, we are selecting a sample mean
    * z=(x̅−μ)/(σ/√n)
    * n = 10
    * z = -1.9309
    * On the TI-84: 1 - normalcdf(-1E99, -1.9309, 0, 1) = 0.9733

```{r}
# Calculating a z score
z <- (60.5 - mu) / (sdev / sqrt(10))
print(z)

#Using the z-score to get the probability
pnorm(z, mean = 0, sd = 1, lower.tail = FALSE)

```


* The Top 13% of all students score higher than ???
    * Since this is the Top 13%, 1 - 0.13 = 0.87
    * Find the z-score corresponding to the area 0.87
    * On the TI-84: invNorm(0.87, 0, 1, LEFT) = 1.1264
    * Use the z-score and the formula: **x=μ+(z ∗ σ)**
    * x = 67.4 + (1.1264 * 11.3) = 80.1283 

*Refer to: Page 227*

```{r}
# Find the z-score based on the given area:
# Area = 1 - 13% = 87%
z <- qnorm(0.13, mean = 0, sd = 1, lower.tail = FALSE)
print(z2)

#Using the z-score to get the value taht the Top 13% of students obtained
x <- mu + (z * sdev)
print(x)

```





